---
title: "Scraping a Wikipedia Table"
output: 
  html_document:
    df_print: paged
---

I tried the rvest package by scraping a Wikipedia table about [communities in the Département "Bas-Rhin" in the region of Alsace in France](https://fr.wikipedia.org/wiki/Liste_des_communes_du_Bas-Rhin). First thing to do is to load the necessary packages: rvest for scraping the website, tidyverse to clean the scraped data.

```{r, results='hide', warning=FALSE, message=FALSE}
library(rvest)
library(tidyverse)
```

Scraping a webpage is surprisingly easy with the convenient rvest package. You only need little knowledge about CSS selectors to access the information you are interested in, in this case the whole table which is converted into a data frame.

```{r, results='hide'}
url <- 'https://fr.wikipedia.org/wiki/Liste_des_communes_du_Bas-Rhin'
webpage <- read_html(url)
table <- html_nodes(webpage,'table.wikitable')
table <- html_table(table, header = TRUE)
bas_rhin <- table %>% 
  bind_rows() %>% 
  as_tibble()
```

The resulting data frame consists of 514 observations and 10 variables.

```{r, echo=FALSE}
bas_rhin
```

A first glance reveals that there are some problems with the data, which is typical for freshly scraped data. The following section deals with all the data preprocessing, for each variable one by one.

```{r, results='hide', warning=FALSE, message=FALSE}
# Nom
head(bas_rhin$Nom)
# Data-Preprocessing: remove additional information
bas_rhin %>% 
  filter(str_detect(Nom, "\\(")) %>% 
  select(Nom)

bas_rhin <- bas_rhin %>% 
  mutate(Nom = str_replace_all(Nom, "\\(préfecture\\)", "")) %>% 
  rename(nom = Nom)

# CodeInsee
head(bas_rhin$CodeInsee)
# Data-Preprocessing: converting to numerical
bas_rhin <- bas_rhin %>% 
  mutate(CodeInsee = as.numeric(CodeInsee)) %>% 
  rename(code_insee = CodeInsee)

# Code postal
head(bas_rhin$`Code postal`)
# Data-preprocessing: split data
# Problem: some have more than one value and got concatenated, Strasbourg even has three
# Strategy: split into several variables
bas_rhin <- bas_rhin %>% 
  mutate(code_postal_1 = str_sub(`Code postal`, 1, 5)) %>% 
  mutate(code_postal_2 = str_sub(`Code postal`, 6, 10)) %>% 
  mutate(code_postal_3 = str_sub(`Code postal`, 11, 15)) %>% 
  select(nom, code_postal_1, code_postal_2, code_postal_3, everything(), -`Code postal`)

# Arrondissement
head(bas_rhin$Arrondissement)
bas_rhin <- bas_rhin %>% 
  rename(arrondissement = Arrondissement)

# Canton
head(bas_rhin$Canton)
bas_rhin <- bas_rhin %>% 
  rename(canton = Canton)

# Intercommunalité
head(bas_rhin$Intercommunalité)
bas_rhin <- bas_rhin %>% 
  rename(intercommunalité = Intercommunalité)

# Superficie
head(bas_rhin$`Superficie(km2)`)
# Data-Preprocessing: converting to numerical
bas_rhin <- bas_rhin %>% 
  mutate(`Superficie(km2)` = str_replace_all(`Superficie(km2)`, ",", ".")) %>% 
  mutate(`Superficie(km2)` = as.numeric(`Superficie(km2)`)) %>% 
  rename(superficie = `Superficie(km2)`)

# Population
head(bas_rhin$`Population(dernière pop. légale)`)
# Data-Preprocessing: removing (2016) and (2015)
bas_rhin <- bas_rhin %>% 
  mutate(`Population(dernière pop. légale)` = str_replace_all(`Population(dernière pop. légale)`, " \\(2016\\)|\\(2015\\)", "")) %>% 
  mutate(`Population(dernière pop. légale)` = str_replace_all(`Population(dernière pop. légale)`, "\\p{WHITE_SPACE}", "")) %>% 
  mutate(`Population(dernière pop. légale)` = as.numeric(`Population(dernière pop. légale)`)) %>%
  rename(population = `Population(dernière pop. légale)`)

# Densité
head(bas_rhin$`Densité(hab./km2)`) 
# Data-Preprocessing: removing white space and converting to numeric
bas_rhin <- bas_rhin %>%   
  mutate(`Densité(hab./km2)` = str_replace_all(`Densité(hab./km2)`, "\\p{WHITE_SPACE}", "")) %>% 
  mutate(`Densité(hab./km2)` = as.numeric(`Densité(hab./km2)`)) %>%
  rename(densite = `Densité(hab./km2)`)

# Delete last column "Modifier"
bas_rhin <- bas_rhin %>% 
  select(-Modifier)
```

Let's take another look at the cleaned data set.

```{r, echo=FALSE}
bas_rhin
```

This data could be used to to create a map of the region, colored by population density.